---
title: "607_Tidyverse_GithubColab"
author: "Fares A"
date: "2024-04-01"
output: html_document
---

Loading a Kaggle dataset on heart failure:
```{r}
df <- read.csv("https://raw.githubusercontent.com/unsecuredAMRAP/HeartFailure_K/main/heart.csv")

colnames(df)
```

Loading the package
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```


## Step 1:

Exploring age and cholesterol levels in relation to heart disease: I will use the mutate, group_by, and the summarise functions here:
```{r}
heart_data <- df %>%
  mutate(AgeGroup = case_when(
    Age < 45 ~ "<45",
    Age >= 45 & Age < 65 ~ "45-65",
    TRUE ~ "65+"
  )) %>%
  group_by(HeartDisease, AgeGroup) %>%
  summarise(AverageCholesterol = mean(Cholesterol, na.rm = TRUE),
            .groups = 'drop')

heart_data
```

Visualize with ggplot2
```{r}
ggplot(heart_data, aes(x = AgeGroup, y = AverageCholesterol, fill = factor(HeartDisease))) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "salmon")) +
  labs(title = "Average Cholesterol Levels by Age Group and Heart Disease",
       x = "Age Group",
       y = "Average Cholesterol Level",
       fill = "Heart Disease") +
  theme_minimal()

```

## Step 2:
 
Now I will use the cor() function to find possible correlations between important variables:
```{r}
correlation_matrix <- df %>%
  select(Age, RestingBP, Cholesterol, MaxHR, Oldpeak) %>%
  cor(use = "complete.obs")
```


Moving on from that, I will try to visualize the pairwise relationships using the GGally::ggpairs function:
```{r}
library(GGally)
df %>%
  select(Age, RestingBP, Cholesterol, MaxHR, Oldpeak) %>%
  ggpairs()

```

===============================================================================

