---
title: "tidyverse_create_assignment"
output: html_document
author: 'Will Berritt'
date: "2024-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective: Choose a dataset from fivethiryeight to demonstrate how the TidyVerse packages can be used to manipulate data. I chose a dataset that measures democratic and republican polling averages since middle of 2018.

```{r, message=FALSE}
library(dplyr)
library(ggplot2)

# Define file path
file_path <- '/Users/williamberritt/Downloads/congress-generic-ballot/generic_ballot_averages.csv'

# Read data from CSV file
political_df <- read.csv(file_path)

# Subset data for Republicans and Democrats
republicans <- subset(political_df, political_df$candidate == 'Republicans')
democrats <- subset(political_df, political_df$candidate == 'Democrats')

# Check dimensions of Democrats and Republicans data
dim(democrats)
dim(republicans)

# Count distinct dates for Democrats and Republicans
n_distinct(democrats$date)
n_distinct(republicans$date)

# Left join Republicans and Democrats data
rep_dem <- left_join(republicans, democrats, by = c('date', 'election', 'cycle'))

# Rename columns for clarity
names(rep_dem) <- c("republicans", "r_ballot_average_estimate", "r_ballot_lo_estimate", "r_ballot_hi_estimate", "date", "election_date", "cycle", "democrats", "d_ballot_average_estimate", "d_ballot_lo_estimate", "d_ballot_hi_estimate")

# Select relevant columns for analysis
clean_df <- rep_dem[c("r_ballot_average_estimate", "d_ballot_average_estimate", "r_ballot_lo_estimate", "d_ballot_lo_estimate", "r_ballot_hi_estimate", "d_ballot_hi_estimate", "date", "election_date", "cycle")]
head(clean_df, 10)

# Convert date columns to Date type
clean_df$date <- as.Date(clean_df$date)
clean_df$election_date <- as.Date(clean_df$election_date)

# Plot Republican and Democrat average estimates over time with election date annotations
ggplot(clean_df, aes(x = date)) +
  geom_line(aes(y = r_ballot_average_estimate, color = "Republican")) +
  geom_line(aes(y = d_ballot_average_estimate, color = "Democrat")) +
  geom_vline(aes(xintercept = as.numeric(election_date)), linetype = "dashed", color = "red") +
  labs(title = "Average Estimates Over Time",
       x = "Date",
       y = "Average Estimate",
       color = "Party") +
  theme_minimal()
```

# Conclusion: Using the dplyr and ggplot packages from TidyVerse, I was able to create a cleaner dataframe that's better prepared for analysis. I was also able to visualize the polling averages over time.
