---
title: "TidyVerse CREATE assignment: forcats vignette"
author: "Naomi Buell"
format: html
editor: visual
---

## Introduction

In this document, I create programming vignette of the **forcats** TidyVerse package. I take [Lana Del Rey's full discography from Kaggle](https://www.kaggle.com/datasets/madroscla/lana-del-rey-genius-discography-released?resource=download) to create an example demonstrating how to use the capabilities of the forcats TidyVerse package for working with factor variables. Factors are used to represent categorical data and with multiple levels.

### Load forcats package

Since forcats is part of the core tidyverse, you can load it with `library(tidyverse)` (or `library(forcats)` to load just forcats on its own). Note that you will need to install these packages before loading.

```{r}
#| label: load packages
#| message: false

# The easiest way to get forcats is to install and load the whole tidyverse:
library(tidyverse)

# Alternatively, just install and load forcats:
library(forcats)
```

## Demo

To begin, load example data for us to work with. All variable types are either character or double variables–no factors yet.

```{r}
#| label: import example data

ldr_raw <-
  read_csv(
    "https://raw.githubusercontent.com/naomibuell/DATA607/main/ldr_discography_released.csv"
  )

head(ldr_raw)
```

### Creating factors

Create factors with the `forcats::as_factor()` function. In the example below, change the variable `category`, which is the categorical classification of an album/song.

```{r}
#| label: assign as factor

# Assign variable as factor:
ldr_cats <- ldr_raw |>
  mutate(category = as_factor(category))

# Category variable is now <fctr> type:
head(ldr_cats)
```

### Browsing Factors

Browse factor levels with the `forcats::levels()`.

```{r}
#| label: get levels

# Browse unique levels of `category` variable:
levels(ldr_cats$category)
```

We see that there are `r length(levels(ldr_cats$category))` levels of the variable `category`. These levels are the titles of studio albums/major EPs, "Non-Album Songs" for promotional singles or songs for soundtracks, and "Other Artist Songs" for songs classified under different artist's names but either feature Lana or are written by Lana.

Get counts of observations in each level in a tibble with `forcats::fct_count`.

```{r}
#| label: get level counts

fct_count(ldr_cats$category)
```

Check if and where level(s) exist in data with `forcats::fct_match`. In this example we verify that her album "Born to Die" is included in the data set and then browse this subset of the data.

```{r}
#| label: check level

# Check matches with "Born to Die"
fct_match(ldr_cats$category, "Born to Die")

# Browse "Born to Die" data
ldr_cats[fct_match(ldr_cats$category, "Born to Die"),]
```

We verified that there are `r fct_match(ldr_cats$category, "Born to Die") |> sum()` songs from that album in this data (this is the length of the bonus track version).

### Reordering levels

Sort levels in by how frequently they appear in the data with `forcats::fct_infreq`.

```{r}
#| label: order levels
levels_ordered <-
  fct_infreq(ldr_cats$category, ordered = NA) |> levels()
levels_ordered
```

Now we can see that most of the songs here are `r levels_ordered[1]` and the album she has with the shortest track list is `r levels_ordered[length(levels_ordered)]`. The following graph confirms this. It is easy to make a bar plot of (ordered) frequencies by factor level:

```{r}
#| label: plot frequencies
ggplot(ldr_cats, aes(y = fct_infreq(category))) +
  geom_bar()
```

### Modifying levels

Use `forcats::fct_lump()` to lump together infrequent levels into an “other” category, specifying the number of levels we want to keep for our factor. In the example below, I narrow the categories down to just the top 5:

```{r}
#| label: lump levels

# Lump levels other than the largest 5 categories
ldr_top5 <- ldr_cats |> 
  mutate(category = fct_lump(category, n = 5))

# Browse top 5 + other category counts in table
top5_counts <- fct_count(ldr_top5$category)
top5_counts
```

Now we see Lana's top 5 longest albums (in terms of track list) above, and there are `r top5_counts[top5_counts$f == "Other", 2]` songs under "other."

## Conclusion

This vignette is part of a collaboration project to build out a book of examples on how to use TidyVerse functions in the our class [TidyVerse GitHub repository](https://github.com/pkowalchuk/SPRING2024TIDYVERSE). View the README.md file to see a description of the commit and a link to the markdown containing this example.
