---
title: "Tidyverse CREATE"
author: "Kaylie Evans"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This project showcases an example of an R Vignette, centered around the powerful visualization capabilities of the ggplot2 package. In this RMD, a dataset sourced from FiveThirtyEight is used, specifically focusing on the age distribution within the U.S. Congress (https://fivethirtyeight.com/features/aging-congress-boomers/). The goal is to demonstrate how to effectively utilize ggplot2, a part of the TidyVerse ecosystem, to create insightful and visually appealing plots from this dataset.

## Load Libraries
```{r load-libraries}
library(tidyverse)
```

## Import CSV into R
```{r import-data}
congress <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/congress-demographics/data_aging_congress.csv")

glimpse(congress)
```

## Essential and Optional Components

The grammar of graphics is the integral of ggplot2. The structure of it takes in data, a coordinate system, and specifies a geom. Depending on the geom, an x and/or y may need to be specified. These are specified in the aes function. 

### Aes values: Example for Histograms

The following are required aesthetics for a histogram: 

- mapping, which is the aes(x = n)

- data, the data set name to be used in the graph


The following are optional aesthetics, followed by their default values:

- stat, default value is "bin"

- position, default value is "stack"

- binwidth, default value is relative to the data set inputs

- bins, default value is 30

- na.rm, default value is FALSE

- orientation, default value is NA

- show.legend, default value is NA

- inherit.aes, default value is TRUE

- aes, the aesthetic field has the ability to contain more aesthetic specifications such as color and shapes in other geoms - default values are null for these additional mappings

### Aes Values Examples with Histograms:
```{r geom-hist-aes}
#histogram of age
ggplot(congress, aes(x = age_years)) + 
  geom_histogram()

#histogram of age with bin number specified
ggplot(congress, aes(x = age_years)) + 
  geom_histogram(bins = 60)

#histogram of age with binwidth specified
ggplot(congress, aes(x = age_years)) + 
  geom_histogram(binwidth = 4)

#histogram of age with bins and colored by generation
ggplot(congress, aes(x = age_years)) + 
  geom_histogram(aes(fill = generation), bins = 20)

```


## Examples of Graphs for Different Variable Types
The below chunks explore different combinations of variable types in ggplot2. The types explored are as follows:

1 Variable:

Continuous = geom_histogram

Discrete = geom_bar


2 Variables: 

Both Continuous = geom_point

1 Continuous, 1 Discrete = geom_boxplot


### One Continuous Variable
```{r geom-hist}
#birthday at start of term histogram
ggplot(congress, aes(x = birthday)) + 
  geom_histogram(bins = 60)
```

### One Continuous Variable with Color Aesthetic 
```{r geom-hist-color}
#age histogram colored by generation
ggplot(congress, aes(x = age_years)) + 
  geom_histogram(aes(fill = generation), bins = 60)
```

### One Discrete Variable  
```{r geom-bar}
#bar chart of generation frequency
ggplot(congress, aes(x = generation)) + 
  geom_bar()
```

### One Discrete Variable with Color Aesthetic 
```{r geom-bar-color}
#bar chart of generation frequency colored by chamber
ggplot(congress, aes(x = generation)) + 
  geom_bar(aes(fill = chamber))
```

### Two Continuous Variables 
```{r geom-point}
#scatter plot of start date and birthday
ggplot(congress, aes(x = start_date, y = birthday)) + 
  geom_point()
```

### Two Continuous Variables with Color Aesthetic 
```{r geom-point-color}
#scatter plot of start date and birthday with color for age to highlight age change over time
ggplot(congress, aes(x = start_date, y = birthday)) + 
  geom_point(aes(color = age_years))
```

### One Continuous Variable and One Discrete Variable
```{r geom-boxplot}
#boxplots for each congress showing age IQR
ggplot(congress, aes(x = congress, y = age_years)) + 
  geom_boxplot(aes(group = congress))
```