---
title: "TIDYVERSE ASSIGNMENT"
author: "Keith DeNivo"
date: "2024-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TidyVerse Vignette

Your task here is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)
Later, you'll be asked to extend an existing vignette.  Using one of your classmate’s examples (as created above), you'll then extend his or her example with additional annotated code. (15 points)
You should clone the provided repository.  Once you have code to submit, you should make a pull request on the shared repository.  You should also update the README.md file with your example.
After you’ve created your vignette, please submit your GitHub handle name in the submission link provided below. This will let your instructor know that your work is ready to be peer-graded.
You should complete your submission on the schedule stated in the course syllabus.

```{r library}
library(tidyverse)
library(httr)
library(dplyr)
library(RCurl)
library(tidyr)
```
## FORCATS VIGNETTE

## forcats
resources:
https://r4ds.hadley.nz/factors.html
https://forcats.tidyverse.org/

forcats is most useful when dealing with factors.

"Factors are used for categorical variables, variables that have a fixed and known set of possible values. They are also useful when you want to display character vectors in a non-alphabetical order."
- https://r4ds.hadley.nz/factors.html

forcats is most useful when dealing with factors.  Factors are character vectors such as eye color.  It works best when there are limited number of character vectors such as eye colors "brown", "blue", "green" etc.

In this car data set there are a couple of columns with character vectors.  We will be looking at the "owner" column in the dataset.  It displays the number of owners a car has, but using characters instead of numerically.

Data source: https://www.kaggle.com/datasets/sukhmandeepsinghbrar/car-price-prediction-dataset/data
 
some forcats functions:
"
fct_reorder(): Reordering a factor by another variable.
fct_infreq(): Reordering a factor by the frequency of values.
fct_relevel(): Changing the order of a factor by hand.
fct_lump(): Collapsing the least/most frequent values of a factor into “other”.
" - https://forcats.tidyverse.org/

We will take a look at fct_reorder() and fct_infreq().

```{r read in data}


car_url <- getURL("https://raw.githubusercontent.com/pkowalchuk/SPRING2024TIDYVERSE/main/cardekho.csv")
car_df <- data.frame(read.csv(text = car_url) )
head(car_df)
names(car_df)

```
##Data
Data set uploaded to the tidyverse git.  Data is read into a dataframe.  The data contains car names, the number of owners the car has had, years, etc.


```{r owners}
# What are the types of owners?  
unique(car_df$owner)
car_df |> count(owner)


owner_levels <- c(
  "Test Drive Car", "First Owner", "Second Owner", "Third Owner", "Fourth & Above Owner" 
)
#setting the order the owners will appear.
```
##Factors

One thing useful about factors is it allows one to set any order of the character vectors, not just in alphabetical order.  Using levels we are about to designate the number of owners in ascending order.
```{r factors}

num_owners <- car_df$owner
o1 <- factor(num_owners, levels = owner_levels)
sort(o1[1:100])
# the values in a column can be sorted by the levels what about the entire dataframe?

car_df <- car_df %>% 
  arrange(factor(owner, levels = owner_levels))
#the dataframe can now be listed by the number of owners.

knitr::kable(car_df[1:20, 1:8], format="html")
```
##Reorder
reordering data frames by factors with user defined levels can be useful, but what about graphs?
```{r graphs }
#another package is ggplot2 for graphs
ggplot(car_df, aes(x = selling_price, y = owner) )+
  geom_boxplot()
# not ordered
ggplot(car_df, aes(x = selling_price, y = fct_reorder(owner, selling_price) ))+
  geom_boxplot()
#ordered by selling price for the factors.
```

##Sorting/Arranging
The factors can be reordered by the which ones had the highest values.


```{r example what forcats is not for}

ggplot(car_df, aes(x = year, y = selling_price, group =year) )+
  geom_boxplot()
  # the sorting with forcats only works with factors and can not be used to order numerical values, there are many other tools to easily sort by numerical value.

```

``` {r by frequency of factor}
car_df |> 
mutate(owner = fct_infreq(owner)) %>%
  ggplot(aes(x = owner)) + 
  ggtitle("how many cars on for each number of owners?")+
  geom_bar() + 
  coord_flip()
#how many of each owner type?
#sort by the frequency of a owner

```
##Conclusion

fct_infreq can be used to sort by the frequency of the factor.
The numerical values for the y values of factors can be reordered using fct_reorder
fct_reorder() and fact_infreq were used to reorder data frames and graphs.