---
title: "Tidyverse CREATE Assignment"
author: "Puja Roy"
date: "3/24/24"
output: openintro::lab_report
---


### Objective

The purpose of this assignment is to create an Example by using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, to create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with the selected dataset.

For this assignment, I selected the coffee shop chain recipes and prices from Kaggle. This dataset contains prices of various coffee drinks from a coffee shop chain in Europe. This is the link to the dataset in Kaggle: https://www.kaggle.com/datasets/deryae0/coffee-shop-chain-recipes-and-prices

### CREATE Vignette

The TidyVerse is a collection of packages designed for data manipulation, visualization, and analysis in R.

```{r load-packages, message=FALSE}
# Load R packages
library(tidyverse)
library(openintro)
library(ggplot2)
```

### readr

**readr**

readr is a fast and efficient package for reading rectangular data files (like CSV or TSV) into R data frames. It provides functions to read data quickly while preserving data types and handling various types of messy data.

I loaded the raw data from my github using the read.csv function. 

- read_csv(), read_excel(), read_table(): Read data from various file formats into data frames.

I explored the data by checking if there were any missing or values. I also printed out summary statistics based on the data.

```{r code-chunk-label}

# Import the dataset
url <- "https://raw.githubusercontent.com/pujaroy280/SPRING2024TIDYVERSE/main/prices.csv"
df_coffee_prize <- read.csv(url)

```

```{r}
# Display the first few rows of the dataset
head(df_coffee_prize)

# Summary statistics
summary(df_coffee_prize)

```

```{r}
num_duplicates <- sum(duplicated(df_coffee_prize))
# Check for duplicates
duplicates <- df_coffee_prize[duplicated(df_coffee_prize), ]
print(duplicates)
```

```{r}
# Print out the column names
print(colnames(df_coffee_prize))
```

### dplyr

**dplyr**

dplyr is a package for data manipulation that provides a set of functions optimized for common data manipulation tasks such as filtering, selecting, mutating, summarizing, and arranging data. It emphasizes a "grammar of data manipulation" approach for easy and intuitive data wrangling.

I removed the last 2 columns using the select function since there weren't any useful info. Then, I removed rows with NA values in any column.

- select(): Select columns based on column names.

```{r}
# Remove the last two columns
clean_df_coffee_prize <- df_coffee_prize %>%
  select(-c(simple, double))

# Remove rows with NA values in any column
clean_df_coffee_prize <- clean_df_coffee_prize %>%
  na.omit()

print(clean_df_coffee_prize)

```

Shown below in the code chunk, I computed the average prices for each combination of drink type and size by using the group_by() function. 

- group_by(): Group data by one or more variables.

In R, the group_by() function is part of the dplyr package, which is a core component of the TidyVerse ecosystem. This function is used for grouping data based on one or more variables, allowing for the application of functions to each group separately.

```{r}
# Compute average prices for each combination of drink type and size
avg_prices <- clean_df_coffee_prize %>%
  group_by(drink_type) %>%
  summarise(
    avg_price_small = mean(price_small, na.rm = TRUE),
    avg_price_medium = mean(price_medium, na.rm = TRUE),
    avg_price_large = mean(price_large, na.rm = TRUE)
  )
print(avg_prices)
```

### ggplot

**ggplot**

To visualize the average prices of each drink type and size based on the average price, I utilized the ggplot library which is a powerful and flexible package for creating static graphics and building complex plots by adding layers and aesthetic mappings.

The ggplot below generates heatmaps to display the average prices of each drink type and size based on the average price.

- ggplot(): Initialize a ggplot object.
- geom_point(), geom_line(), geom_bar(): Add geometric objects like points, lines, or bars to the plot.
- aes(): Define aesthetic mappings.
- labs(): Add titles, labels, and captions to the plot.
- scale_fill_gradient(), scale_color_manual(): Adjust color scales.


```{r}
# Create heatmap with reversed axes
ggplot(avg_prices, aes(x = avg_price_small, y = drink_type, fill = avg_price_small)) +
  geom_tile(color = "darkgrey") +
  labs(title = "Average Small Coffee Prices by Drink Type", x = "Average Price (Small)", y = "Drink Type", fill = "Average Price") +
  scale_fill_gradient(low = "darkgrey", high = "brown")

```

```{r}
# Create heatmap with reversed axes
ggplot(avg_prices, aes(x = avg_price_medium, y = drink_type, fill = avg_price_medium)) +
  geom_tile(color = "darkgrey") +
  labs(title = "Average Medium Coffee Prices by Drink Type", x = "Average Price (Medium)", y = "Drink Type", fill = "Average Price") +
  scale_fill_gradient(low = "darkgrey", high = "brown")
```

```{r}
# Create heatmap with reversed axes
ggplot(avg_prices, aes(x = avg_price_large, y = drink_type, fill = avg_price_large)) +
  geom_tile(color = "darkgrey") +
  labs(title = "Average Large Coffee Prices by Drink Type", x = "Average Price (Large)", y = "Drink Type", fill = "Average Price") +
  scale_fill_gradient(low = "darkgrey", high = "brown")
```
















...

