---
title: "HW10"
author: "Andrianina David Raharijao"
date: "2024-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
```

```{r tiddyverse package, message = FALSE, warning=FALSE}
library("tidyverse")
library("tidytext")
library("janeaustenr")
library("tidyr")
library("dplyr")
library("stringr")
library("ggplot2")
```

## Web API

First get the data using API:

```{r q1}
api_url <- "https://api.nytimes.com/svc/topstories/v2/us.json?api-key=XyD2Prjgvsw5Psu7IYBatJGEDtrJnIfw"
response <- GET(api_url)
parsed_content <- content(response, "text")
df <- fromJSON(parsed_content)
s_df <- df$results
```

##Dplyr

Filter articles with the word "Trump" in the title
```{r q2}
filtered_articles <- s_df %>% 
  filter(grepl("Trump", title, ignore.case = TRUE))
```

Arrange articles by publication date
```{r q3}
sorted_articles <- s_df %>% 
  arrange(desc(published_date))
```

##ggplot2

Bar chart of articles per item type
```{r q4}
s_df %>% 
  count(item_type) %>% 
  ggplot(aes(x = item_type, y = n)) + 
  geom_bar(stat = "identity")
```

Time series graph for articles over time
```{r q5}
s_df %>% 
  mutate(published_date = as.Date(published_date)) %>% 
  count(published_date) %>% 
  ggplot(aes(x = published_date, y = n)) + 
  geom_line()
```

## Extended

## Bar chart of articles per item type

```{r bar-chart}
s_df %>% 
  count(item_type) %>% 
  ggplot(aes(x = item_type, y = n)) + 
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Number of Articles by Item Type", x = "Item Type", y = "Count") +
  theme_minimal()
```

## Group by Section: Get the number of articles per section

```{r}
articles_by_section <- s_df %>% 
  group_by(section) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))
articles_by_section
```

## Summarize: Count the number of articles, average title length, etc.
```{r}
summary_stats <- s_df %>% 
  summarize(
    total_articles = n(),
    avg_title_length = mean(nchar(title)),
    avg_abstract_length = mean(nchar(abstract))
  )

summary_stats
```

## Filter by Date Range: Focus on articles published in the last month
```{r}
last_month <- Sys.Date() - 30
recent_articles <- s_df %>% 
  mutate(published_date = as.Date(published_date)) %>% 
  filter(published_date >= last_month)

head(recent_articles)
```

# Basic Sentiment Analysis: Analyze sentiment of titles
```{r}
bing <- get_sentiments("bing")
title_sentiment <- s_df %>% 
  unnest_tokens(word, title) %>% 
  inner_join(bing, by = "word") %>% 
  count(sentiment) %>% 
  ggplot(aes(x = sentiment, y = n, fill = sentiment)) + 
  geom_col() + 
  labs(title = "Sentiment Analysis of Article Titles", x = "Sentiment", y = "Count") +
  theme_minimal()

title_sentiment
```
