---
title: "Vignette for Purrr"
author: "Rashad Long"
format: html  
editor: visual
---

## Introduction

The `purrr` package provides functions that eliminate the need for many common for loops. They are more consistent and thus easier to learn than many of the alternative functions in the base R package. It allows you to generalize a solution to every element in a list. It also allows you get lots of small pieces and compose them together with the pipe.

```{r}
#| warning: FALSE

# Load the purr package along with other tidyverse packages
if(!require("tidyverse")) {install.packages("tidyverse"); library("tidyverse")}


```

## Map Functions

The `purr` package provides functions that allows you to loop over a vector, do something to each element, and save the results

-   `map()` makes a list.

-   `map_lgl()` makes a logical vector.

-   `map_int()` makes an integer vector.

-   `map_dbl()` makes a double vector.

-   `map_chr()` makes a character vector.

The map functions make it easier to write and to read than for loops.

We will import a dataset from Kaggle that contains information about NFL passing statistics from players since 2001.

```{r}

# Lets import the passing_cleaned csv
passing_stats <- read.csv("passing_cleaned.csv")
passing_stats <- as_tibble(passing_stats)

# Lets extract the variables Age, G, Yds, TD, Cmp, Int, Rate
passing_stats <- passing_stats |> 
  select(Age, G, Yds, TD, Cmp, Int, Rate) |> 
  arrange(desc(Yds))

passing_stats

```

Imagine if we wanted to compute the mean, median and sd for every column. We can use the map function to do this. Since they will generate doubles we need to us `map_dbl()`.

```{r}

# Compute mean for every column
passing_stats_mean <- passing_stats |> 
  map_dbl(mean, na.rm = TRUE)

# Compute median for every column
passing_stats_median <- passing_stats |> 
  map_dbl(median, na.rm = TRUE)

# Compute sd for every column
passing_stats_sd <- passing_stats |>
  map_dbl(sd, na.rm = TRUE)

passing_stats_mean
passing_stats_median
passing_stats_sd

```

As you can see the map function allowed us to apply some solution to an entire data-frame in one command!

## Extend - Semyon Toybis

I am interested in comparing the amount of total career touchdowns over the course of the careers of some notable quarterbacks.

I start by reading in the original file to have the player names and create a vector of quarterbacks that I am interested in comparing. I subset the original data frame for the QBs I am interested in.

```{r}
statsWithNames <- read.csv("passing_cleaned.csv")

qbsOfInterest <- c('Eli Manning', 'Ben Roethlisberger', 'Aaron Rodgers', 'Joe Burrow', 'Lamar Jackson', 'Patrick Mahomes')

statsFiltered <- subset(statsWithNames, Player %in% qbsOfInterest)

head(statsFiltered, n=10)
```

Next, I create a new data frame that has the names of the players, their year in the league, and the TDs for that year. In that new data frame, I create a column that has the cumulative sum of TDs for each player. I then sort the data frame by Player and Year, group by player, and add a column that has the year in the league for the player, which is the row number. This allows us to analyze how many TDs each QB had in their first, second, third, etc. year in the league. I used the Dplyr functions 'select', 'group_by', and 'mutate'.

```{r}

cumulativeTDs <- statsFiltered |> select(Player, TD, Year) |> group_by(Player) |>
  mutate(cumualativeTDs = cumsum(TD))

cumulativeTDs <- cumulativeTDs |> arrange(Player, Year) |>
  group_by(Player) |> mutate(YearInLeage = row_number())


head(cumulativeTDs, n = 10)
```

Below I visualize the cumulative TDs over time by player. I use the ggplot2 package.

```{r}
cumulativeTDs |> ggplot(aes(x = YearInLeage, y = cumualativeTDs, color = Player)) + geom_line() + ggtitle('Amount of career touchdowns over time by QB')
```

Interestingly, Patrick Mahomes is by far ahead of the other QBs for total career touch downs at this point in his career. Aaron Rodgers, who has the most TDs of the group interestingly spent a few years in the league before he recorded his first TD (those familiar with the NFL will know it was because he was a backup for Brett Favre and didn't play his first few years). It would be interesting to include QBs like Peyton Manning and Tom Brady in this analysis, but their careers started before 2001 so the data set does not fully capture their careers.
